% \paragraph{Caption is the most important thing; image next. Paragraph is the least important---even our INLG paper said so! (CITATION). \cite{huang-etal-2023-summaries}}
% \paragraph{Caption is the most important thing; image next. Paragraph is the least important---even our INLG paper said so! (CITATION). \cite{huang-etal-2023-summaries}}

% \input{figures/fig01/fig01_ablation_study}
%\kenneth{TODO Human Evaluation here: https://openreview.net/forum?id=ghxUYlolwr\#discussion}


%\input{table/table_human rank}
\input{figures/fig_humanRankDist/humanRankDist}

%\kenneth{I'd use All-Profile, No-Profile instead of All-profile, No-profile. Use upper case P, as they are basically names of conditions.}

%\kenneth{TODO: Which model do we use in human eval????}

%\sam{ADD-START}
%To supplement automatic metrics, which do not always correlate with quality, we conducted a human evaluation.
We recruited 10 US-based NLP researchers (PhD students with publication or review experience) for our human evaluation.
%\kenneth{Be more specific, are they PhD students? or?}\sam{corrected!} 
The study involved ranking captions for 50 figures randomly selected from the arXiv \textit{cs.CL} domain. 
For each figure, we provided the figure's image, the paper's title, and abstract as context. 
Participants then ranked four corresponding captions based on how well each one helped them understand the figure. The four captions were:
%\kenneth{What exactly is the QUESTION we asked the participants? Don't just say QUALITY.} \sam{corrected! Also included the full instruction in the Appendix as requested in the EMNLP checklist}
{\em (i)} the original \textit{Gold} caption, and
{\em (ii)} \textit{No-Profile}, 
{\em (iii)} \textit{1-Profile}, and 
{\em (iv)} \textit{All-Profile} settings generated by our best-performing model (\textbf{Gemini}, based on automatic metrics). 
See \autoref{app:human-eval} for details.
%The complete experimental setup, which follows prior work~\cite{hsu2025large}, is detailed in \autoref{app:human-eval}.
%with the complete experimental setup and participant instructions detailed in \autoref{app:human-eval}.
%The ranking results are presented in 

\paragraph{1-Profile was the most preferred condition.} Our human evaluation result shows a clear preference for \textit{1-Profile} setting, which achieved the best average rank of 2.27. The other models followed in order: \textit{No-Profile} (2.48) and \textit{All-Profile} (2.54). Interestingly, the author-written \textit{Gold} captions were ranked last overall with an average rank of 2.71. Such outcome is reflected in the preference distribution (\autoref{fig:human-rank}): The \textit{1-Profile} configuration received the most first-place votes (30.6\%) and the fewest last-place votes (17.8\%). On contrast, the \textit{Gold} captions were ranked last most frequently (31.2\%).
%\kenneth{TODOs: As we do not have a table now, write a few sentences here to report the avg rank for each model here.}\sam{done!}
Post-hoc Friedman-Nemenyi tests showed that 1-Profile and No-Profile captions were significantly preferred over gold ($p < 0.001$ and $p = 0.03$, respectively).

\paragraph{Trade-offs between human-perceived quality and similarity to gold captions.} 
While the All-Profile setting generated captions with higher reference similarity (\autoref{tab:llm_comparison}), human judges significantly preferred captions from 1-Profile ($p = 0.006$). 
This suggests that optimizing for similarity may also reproducing flaws from the inconsistently quality's caption from arXiv reference~\cite{huang-etal-2023-summaries}, reducing the perceived quality.

%This may stem from the variable quality of the reference arXiv captions themselves~\cite{huang-etal-2023-summaries}. 
%Consequently, when a model too closely mimics the gold standard, it may inherit stylistic flaws, causing perceived quality to decline as similarity increases excessively.

%\sam{ADD-END}
%------------ dead kitten ----------
\begin{comment}


    %While the All-Profile setting generated captions with higher reference similarity, with higher similarity to the references (\autoref{tab:llm_comparison}), human judges significantly preferred captions from 1-Profile ($p = 0.006$). 

    % This likely reflects the inconsistent quality of arXiv captions~\cite{huang-etal-2023-summaries}.
Models that closely mimic these references risk reproducing their flaws, reducing perceived quality as similarity increases.
\end{comment}



